---
title: "Exploratory Data Analysis in R"
output: html_notebook
---

## Set locale

```{r}
Sys.setlocale(category = 'LC_TIME', locale = 'en_US.UTF-8')
```

```{r}
install.packages("visdat")
```

## Load libraries

```{r}
library(ggplot2)
library(visdat)
library(dplyr, quietly = T)
library(tidyr)
```


## Read and prepare data

```{r}
df <- read.csv('electricity.csv', stringsAsFactors = FALSE)
```


## Take a look at the data

```{r}
head(df)
```


## Check data structure

```{r}
str(df)
```


## Check data summary

```{r}
summary(df)
```


## Visualize data structure

```{r}
vis_dat(df)
```


## What questions we can ask?

- How consumption is distributed by day, week, month, season, year?
- What is the structure of electricity production? Which energy source contributes the most?
- What is the difference between production and consumption? Is there a deficit or proficit?
- Can we recreate typical work day with this data?
- Additional - is there relation between temperature and electricity consumption?

## How do we have to prepare data to answer those questions?

- Convert 'season' to a ordered factor
- Extract date part from the datetime variable
- Extract weekday from the date
- Reshape data

## Prepare data

```{r}
df <- df %>% 
  mutate(date  = as.Date(datetime),
         wday = weekdays(date, abbreviate = TRUE),
         wday = factor(wday, ordered = TRUE, 
                       levels = c('Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat', 'Sun')),
         season = factor(season, ordered = TRUE,
                         levels = c('Winter', 'Spring', 'Summer', 'Autumn')))
```



## Visualise distribution of electricity consumption (by season, month, weeekday)

```{r}
ggplot(df, aes(x=consumption)) + geom_histogram(start="count")
```


## Compare electricity consumption (by season, month, weekday)

```{r}

```

## Visualize electricity consumption change in time

```{r}
ggplot(df, aes(x=date, y=consumption)) + geom_point(size=0.2)
```

## Visualize typical day in terms of electricity consumption

```{r}

```


## Compare amount of electricity produced and consumed

```{r}

```



## Reshape data

```{r}
df_long <- df %>% 
  select(date, wday, hour, month, season, nuclear:hydro_gaes) %>% 
  gather(key = 'type', value = 'production', 
         -date, -wday, -hour, -month, -season)
```

```{r}
head(df_long)
```

## Compare production by source type

```{r}
df_long %>% group_by(type) %>% summarise(production=sum(production)) %>% 
  ggplot(aes(x=reorder(type, production), y=production)) + geom_col()
```

```{r}
df_long %>% group_by(date, type)  %>% summarise(production = sum(production)) %>%
  ggplot(aes(x=date, y=production, fill=type)) + geom_area()
```


## Load additional data

```{r}
weather <- read.csv('weather_vs_electricity_consumption.csv', stringsAsFactors = FALSE)
```

## Preview additional data

```{r}
head(weather)
```

## Convert season variable to ordered factor

```{r}
weather <- weather %>% 
  mutate(season = factor(season, ordered = TRUE,
                         levels = c('Winter', 'Spring', 'Summer', 'Autumn')))
```

## Visualise relation between temperature (tavg) and electricity consumption (consumption)

```{r}

```

