---
title: "Problem 5"
output: html_notebook
---


```{r}
ggplotRegression <- function (fit, i) {

require(ggplot2)
ggplot(fit$model, aes_string(x = names(fit$model)[i], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[i]], 5),
                     " P =",signif(summary(fit)$coef[i,4], 5)))
}
```

```{r}
y <- cars$speed
yu <- y + 10
yl <- y - 10
x <- cars$dist
xu <- x + 10
```

```{r}

fit <- lm(y ~ x)
fit_up <- lm(yu ~ x)
fit_down <- lm(yl ~ x)
fit_right <- lm(y ~ xu)
```
#### A

When we add or subtract 10 from X or Y we only affect the intercept term. Neither R squared nor slope is changing in this case. This is constant change(shift) that can be described by intercept term fully. 


```{r}
ggplot(data.frame(x=x, y=y)) + geom_point(aes(x=x, y=y))
```

```{r}
ggplotRegression(fit, 2)
```
```{r}
ggplotRegression(fit_up, 2)
```
```{r}
ggplotRegression(fit_down, 2)
```

#### B 

Same for chaning x as described in **A**

```{r}
ggplotRegression(fit_right, 2)
```

#### C
As we can see - scaling doesn't affect value of R2. Also, we can see that intercept value is significantly lower and very close to 0 in this case (because we removed unconidional mean). Also, Slope is much bigger. 

Shifting the regression by 10 have same effect as without demeaning - intercept is changed by small value (because 10 is relatively small value on the Y initial range)
```{r}
fit <- lm(y ~ x)
fit_norm <- lm(scale(y) ~ scale(x))
fityu_norm <- lm(scale(yu) ~ scale(x))
```



```{r}
ggplotRegression(fit_norm, 2)
```

```{r}
ggplotRegression(fityu_norm, 2)
```






